---
title: "Titanic Survival Analysis"
author: "Jirawan Phromtawepong"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    theme: cosmo
---

# 🛳 Introduction

This project explores survival patterns in the Titanic dataset using R. The analysis focuses on how **gender**, **passenger class**, and **family presence** influenced the likelihood of survival.

The goal is to demonstrate end-to-end data analysis skills: from data cleaning and transformation to visualization and insight generation. This project is part of my data portfolio to showcase my analytical thinking and ability to communicate findings.

---

# 📦 Dataset Overview

```{r load-libraries, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
```

```{r load-data}
titanic <- read_csv("titanic.csv")
```

```{r initial-explore}
glimpse(titanic)
summary(titanic)
```

---

# 🧹 Data Cleaning & Preparation

```{r cleaning}
titanic_clean <- titanic %>%
  filter(!is.na(Age)) %>%
  mutate(
    FamilySize = SibSp + Parch,
    TravelAlone = ifelse(FamilySize == 0, "Alone", "WithFamily"),
    Survived = factor(Survived, levels = c(0,1), labels = c("Died", "Survived")),
    Pclass = factor(Pclass),
    Sex = factor(Sex)
  )
```

---

# 📊 Analysis & Visualization

## 1. Survival Rate by Gender

```{r gender-plot}
ggplot(titanic_clean, aes(x = Sex, fill = Survived)) +
  geom_bar(position = "fill") +
  labs(title = "Survival Rate by Gender", y = "Proportion")
```

**Insight:**

Female passengers had a **significantly higher survival rate** compared to male passengers. Over 70% of females survived, versus less than 20% of males. This aligns with the evacuation protocol of prioritizing "women and children first."

---

## 2. Survival Rate by Passenger Class

```{r class-plot}
ggplot(titanic_clean, aes(x = Pclass, fill = Survived)) +
  geom_bar(position = "fill") +
  labs(title = "Survival Rate by Passenger Class", y = "Proportion")
```

**Insight:**

Passengers in **1st class** had the highest survival rate, followed by 2nd class, while **3rd class** passengers had the lowest. This highlights a likely influence of **socioeconomic status** and cabin accessibility during evacuation.

---

## 3. Survival Rate: Traveling Alone vs With Family

```{r family-plot}
ggplot(titanic_clean, aes(x = TravelAlone, fill = Survived)) +
  geom_bar(position = "fill") +
  labs(title = "Survival Rate: Alone vs With Family", y = "Proportion")
```

**Insight:**

Passengers traveling **with family** had a significantly higher survival rate than those traveling alone. This may reflect both practical support and social prioritization for groups during evacuation.

---

# 🛠 Tools Used

- **Language**: R
- **Libraries**: `readr`, `dplyr`, `ggplot2`
- **Skills Demonstrated**:
  - Data cleaning
  - Feature engineering
  - Exploratory data analysis
  - Data visualization
  - Insight communication

---


```text
titanic-survival-analysis/
├── data/
│   └── titanic.csv
├── R/
│   └── titanic_analysis.Rmd
├── outputs/
│   └── titanic_analysis.html  # (after knitting)
├── README.md
└── LICENSE
```
